---
title: "Compute a whole family"
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

You rarely want one variable. You want every low-flow indicator, or every
temperature indicator your data can feed. The collection is queryable, so
that is two lines rather than a list you maintain by hand.

```{r setup}
library(card4r)
library(airGRdatasets)
```

## The data

[airGRdatasets](https://cran.r-project.org/package=airGRdatasets) ships
daily records for gauged French catchments, which is enough to run
without downloading anything. Three of them, chosen for contrast:

```{r}
ids <- c("A273011002", "X031001001", "Y862000101")
do.call(rbind, lapply(ids, function(i) {
  d <- get(i)
  data.frame(id = i, name = d$Meta$Name, area_km2 = round(d$Meta$Area))
}))
```

card reads its input columns by name, so the only preparation is to give
them the names the cards use: `Q` for discharge in m³·s⁻¹, `R` for
precipitation, `T` for temperature, `ETP` for potential
evapotranspiration. `card_list()` and `card_info()` state what each card
needs.

```{r}
ts <- do.call(rbind, lapply(ids, function(i) {
  d <- get(i)$TS
  data.frame(id = i, date = as.Date(d$Date),
             R = d$Ptot, T = d$Temp, ETP = d$Evap,
             Q = d$Qls / 1000)          # airGRdatasets serves L/s
}))
str(ts)
```

## Select a family, then compute it

Ask the collection for a family. Facets accept English or French, and the
[catalogue](https://lou-heraut.github.io/card/CARDS) lists the values each
one takes.

```{r}
temp <- card_list(domain = "temperature")
nrow(temp)
```

Not all of those run on a temperature series alone: some compare
discharge to temperature. You do not have to know which, the listing says
so.

```{r}
table(temp$input_vars)
```

```{r}
temp <- temp[temp$input_vars == "T", ]
nrow(temp)
```

One row per **variable**. What you extract is a **card**, and one card can
produce several variables: the twelve `mean-TMA_*` are the monthly columns
of the single card `mean-TMA_month`. The `card` column holds that name.

```{r}
unique(temp$card)
```

That is the whole gesture:

```{r}
res <- card_extract(ts[, c("id", "date", "T")], cards = unique(temp$card))
res$data[["mean-TA"]]
```

Which reads as a catchment gradient rather than as three numbers: the
Bruche drains the Vosges foothills, the Durance at Embrun is alpine, and
the Taravo is Corsican.

The metadata comes with it, one row per variable produced, saying which
card each one came from:

```{r}
head(res$meta[, c("card", "variable_en", "unit_en", "name_en")], 3)
```

## The same two lines, anywhere in the collection

Nothing above is specific to temperature. Swap the facet and the input
column:

```{r}
lf <- card_list(phenomenon = "low flows", output = "series")
lf <- lf[lf$input_vars == "Q", ]
c(variables = nrow(lf), cards = length(unique(lf$card)))
```

```{r}
res <- card_extract(ts[, c("id", "date", "Q")], cards = unique(lf$card))
head(res$data[["VCN10"]], 3)
```

`card_trend()` then tests every one of them, in one call:

```{r}
tr <- card_trend(res)
tr$data[["VCN10"]][, c("id", "h", "p", "a")]
```

`h` is significance at the requested level, `0.1` by default, and `a` the
Sen slope per year. None of these three is significant over twenty years,
which is the honest answer for a record that short.

## Where to look next

This page shows a gesture, not an inventory. The variables themselves are
published as a catalogue, in
[English](https://lou-heraut.github.io/card/CARDS) and in
[French](https://lou-heraut.github.io/card/CARDS.fr), generated from the
cards so that it cannot drift from them. `card_info("VCN10")` draws a
single card, step by step, in your console.
